Supplement to ``Local clustering of density upper level sets''

### 2019-01-22 icml/icml_local_density_clustering_supplement.tex
added todays changes

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{xr-hyper}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}

@@

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\TV}{\mathrm{TV}}

@@

%%% Tilde notation for quantities over the expansion set 
\newcommand{\wn}{\widetilde{n}}
\newcommand{\wX}{\widetilde{\Xbf}}
\newcommand{\wx}{\widetilde{x}}
\newcommand{\wz}{\widetilde{z}}
\newcommand{\wbz}{\widetilde{\bf{z}}}
\newcommand{\wu}{\widetilde{u}}
\newcommand{\wPbb}{\widetilde{\Pbb}}

@@

In this supplement, we present proofs for ``Local Clustering of Density Upper Level Sets''. We begin by providing technical lemmas, before moving on to proving the main results of the paper.

Throughout, we will fix $\Aset \subset \Rd$ to be an arbitrary set. To simplify expressions, for the $\sigma$-expansion $\Asig$, we will write the set difference between $\Asig$ and the $(\sigma + r)$-expansion $\Aset_{\sigma + r}$ as 
\begin{equation*}
\Asigr := \set{x: 0 < \rho(x, \Asig) \leq r},
\end{equation*}
where $\rho(x, \Aset) = \min_{x' \in \Aset} \norm{x - x'}$.

@@

\subsection{Technical Lemmas}

@@

\begin{lemma}
	\label{lem: expansion_sets}
	For any $\sigma > 0$ and the $\sigma$-expansion $\Asig = \Aset + \sigma B$, 
	\begin{equation*}
	\sigma B \subset \Asig, ~~\mathrm{and~ }\nu(\Aset + \sigma B) \leq \nu((1 + \sigma)\Aset) = (1 + \sigma)^d \nu(\Aset).
	\end{equation*}
\end{lemma}

We will need to carefully control the volume of the expansion set using the above estimate; Lemma \ref{lem: Taylor_series} serves this purpose.
\begin{lemma}
	\label{lem: Taylor_series}
	For any $0 \leq x \leq 1/2d$,
	\begin{equation*}
	(1 + x)^d \leq 1 + 2dx.
	\end{equation*}
\end{lemma}
The proof of Lemma \ref{lem: Taylor_series} is based on approximation via Taylor series, and we omit it.

@@

The proof of Theorem \ref{thm: inverse_mixing_time_lower_bound} also depends on a parameter -- which we term \emph{discrete local spread} -- to handle the mixing over very small steps. Formally, the discrete local spread $\pi_1(G)$ is given by
\begin{equation}
\label{eqn: local_spread}
\pi_1(G) := \frac{d_{\min}(G)^2}{10 \vol(V; G)} 
\end{equation}
where $d_{\min}(G) = \min_{v \in V} d(v)$ is the minimum degree in $G$. Intuitively, the discrete local spread gauges how much the walk given by $\Wbf$ has mixed after one step, starting from any node $v$. We will denote $\pi_1(G_{n,r}[\Csig[\Xbf]])$ by $\widetilde{\pi}_{1,n}$.

\textcolor{red}{Prove Lemma \ref{lem: local_spread_lower_bound}}.

\subsection{Cut and volume estimates}
\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $r < \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq 2 \nu(\Csig) \frac{rd}{\sigma}  \left(\lambda_{\sigma} - \frac{r^{\gamma}}{\gamma + 1}\right)
	\end{equation*}	
\end{lemma}
\begin{proof}
	Recalling that $f$ is the density function for $\Pbb$, we have
	\begin{equation}
	\label{eqn: integral_over_epsilon_neighborhood}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) dx
	\end{equation}
	We partition $\Csigr$ into slices, based on distance from $\Csig$, as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Rd: t_{i,k} < \frac{\rho(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k}
	\end{equation*}
	where $t_i = i/k$ for $i = 0, \ldots, k - 1$. As a result,
	\begin{equation*}
	\int_{\Csigr} f(x) dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation*}
	We substitute
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) := \nu_{i+1,k} - \nu_{i,k}. 
	\end{equation*}
	where for simplicity we've written $\nu_{i,k} = \nu(\Csig + rt_{i+1,k}B)$.
	This, in concert with the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - (rt_{i,k})^{\gamma},
	\end{equation*}
	which follows from \ref{asmp: bounded_density} and \ref{asmp: low_noise_density}, yields
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - (rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \sum_{i = 1}^{k} 
	\underbrace{\nu_{i,k} \biggl( \left[\lambda_{\sigma} - (rt_{i,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - (rt_{i-1,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi_k}
	\end{align}

@@

\subsection{Proof of Theorem \ref{thm: conductance_upper_bound}}
The proof of Theorem \ref{thm: conductance_upper_bound} is more or less given by Lemmas \ref{lem: expected_density_cut}, \ref{lem: expected_density_volume}, \ref{lem: prob_bound_cut}, and \ref{lem: prob_bound_vol}. All that remains is some algebra, which we take care of below.

@@

>>>>

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{xr-hyper}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}

@@

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Id}{\mathrm{Id}}

@@

%%% Tilde notation for quantities over the expansion set 
\newcommand{\wn}{\widetilde{n}}
\newcommand{\wX}{\widetilde{\Xbf}}
\newcommand{\wx}{\widetilde{x}}
\newcommand{\wz}{\widetilde{z}}
\newcommand{\wbz}{\widetilde{\bf{z}}}
\newcommand{\wu}{\widetilde{u}}
\newcommand{\wPbb}{\widetilde{\Pbb}}
\newcommand{\wf}{\widetilde{f}}

@@

In this supplement, we present proofs for ``Local Clustering of Density Upper Level Sets''. Sections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. \ref{sec: mixing_time_on_graphs} develops a bound of the form of \eqref{eqn: average_conductance}, which we recall links the conductance function to mixing time; this will be necessary for both Theorems \ref{thm: inverse_mixing_time_lower_bound_nonconvex} and \ref{thm: inverse_mixing_time_lower_bound}. \ref{sec: non_convex_conductance_function_and_local_spread} and \ref{sec: inverse_mixing_time_lower_bound_nonconvex} give the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex}, while \ref{sec: convex_population_conductance_function}- \ref{sec: inverse_mixing_time_lower_bound} give the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound}.

\subsection{Volume estimates}
\label{sec: volume_estimates}

We will fix $\Aset \subset \Rd$ to be an arbitrary set. To simplify expressions, for the $\sigma$-expansion $\Asig$, we will write the set difference between $\Asig$ and the $(\sigma + r)$-expansion $\Aset_{\sigma + r}$ as 
\begin{equation*}
\Asigr := \set{x: 0 < \dist(x, \Asig) \leq r},
\end{equation*}
where as a reminder $\dist(x, \Aset) = \min_{x' \in \Aset} \norm{x - x'}$.

@@

@@

\begin{lemma}
	\label{lem: expansion_sets}
	For any $\delta > 0$ and $x \in \Aset_{\sigma}$,
	\begin{equation*}
	\sigma B \subset \Asig, ~~\mathrm{and~ }\nu(\Asig + \delta B) \leq \nu\left(\left[1 + \frac{\delta}{\sigma}\right]\Asig\right) = \left(1 + \frac{\delta}{\sigma}\right)^d \nu(\Asig)
	\end{equation*}
	where $\Asig = \Aset + \sigma B$ is the $\sigma$-expansion of $\Aset$.
\end{lemma}

We will need to carefully control the volume of the expansion set using the above estimate; Lemma \ref{lem: Taylor_series} serves this purpose.
\begin{lemma}
	\label{lem: Taylor_series}
	For any $0 \leq x \leq 1/2d$,
	\begin{equation*}
	(1 + x)^d \leq 1 + 2dx.
	\end{equation*}
\end{lemma}
\begin{proof}
	We take the binomial expansion of $(1 + x)^d$:
	\begin{align*}
	(1 + x)^d & = \sum_{k = 0}^{d} {d \choose k} x^k \\
	& = 1 + dx + dx\left(\sum_{k = 2}^d \frac{{d \choose k} x^{k - 1}}{d}\right)\\
	& \leq 1 + dx + dx\left(\sum_{k = 2}^d \frac{d^{k}}{(2d)^{k-1}d}\right) \\
	& \leq 1 + 2dx.
	\end{align*}
\end{proof}

@@

\subsection{Density-weighted cut and volume estimates}
\label{sec: density_weighted_cut_and_volume_estimates}

\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $r < \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq 2 \nu(\Csig) \frac{rd}{\sigma}  \left(\lambda_{\sigma} - \frac{r^{\gamma}}{\gamma + 1}\right)
	\end{equation*}	
\end{lemma}
\begin{proof}
	Recalling that $f$ is the density function for $\Pbb$, we have
	\begin{equation}
	\label{eqn: integral_over_epsilon_neighborhood}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) dx
	\end{equation}
	We partition $\Csigr$ into slices, based on distance from $\Csig$, as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Rd: t_{i,k} < \frac{\dist(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k}
	\end{equation*}
	where $t_i = i/k$ for $i = 0, \ldots, k - 1$. As a result,
	\begin{equation*}
	\int_{\Csigr} f(x) dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation*}
	We substitute
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) := \nu_{i+1,k} - \nu_{i,k}. 
	\end{equation*}
	where for simplicity we've written $\nu_{i,k} = \nu(\Csig + rt_{i+1,k}B)$.
	This, in concert with the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - (rt_{i,k})^{\gamma},
	\end{equation*}
	which follows from \ref{asmp: bounded_density} and \ref{asmp: low_noise_density}, yields
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - (rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \sum_{i = 1}^{k} 
	\underbrace{\nu_{i,k} \biggl( \left[\lambda_{\sigma} - (rt_{i,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - (rt_{i-1,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi_k}
	\end{align}

@@

\subsection{Proof of Theorem \ref{thm: conductance_upper_bound}}
\label{sec: proof_of_theorem_1}

The proof of Theorem \ref{thm: conductance_upper_bound} is more or less given by Lemmas \ref{lem: expected_density_cut}, \ref{lem: expected_density_volume}, \ref{lem: prob_bound_cut}, and \ref{lem: prob_bound_vol}. All that remains is some algebra, which we take care of below.

@@

\subsection{Mixing time on graphs}
\label{sec: mixing_time_on_graphs}

For $N \in \mathbb{N}$ and a set $V$ of $N$ vertices, take $G = (V,E)$ to be an undirected and unweighted graph, with associated adjacency matrix $\Abf$, random walk matrix $\Wbf$, and stationary distribution $\pibf = (\pi_u)_{u \in V}$ where $\pi_v = \frac{\Dbf_{vv}}{\vol(V; G)}$. For $v \in V$, 
\begin{equation}
\label{eqn: random_walk}
q_{vu}^{(m)} = e_v\Wbf^m e_u, ~~ \qbf_{v}^{(m)} = \left(q_{vu}^{(m)}\right)_{u \in V}, ~~ \qbf_v = (\qbf^{(1)}_{v\cdot},\qbf^{(2)}_{v\cdot}, \ldots), 
\end{equation}
denote respectively the $m$-step transition probability, distribution, and sequence distributions of the random walk over $G$ originating at $v$. Letting $\qbf = (\qbf_v)_{v \in V}$, the relative pointwise mixing time is thus
\begin{equation*}
\tau_{\infty}(\qbf; G) = \min\set{m: \forall u,v \in V, \frac{\abs{q_{vu}^{(m)} - \pibf_u}}{\pibf_u} \leq 1/4} 
\end{equation*}

The local spread facilitates conversion between $\tau_{\infty}(\qbf_v; G)$ and the more easily manageable \emph{total variation} mixing time, given by
\begin{equation*}
\tau_1(\rhobf; G) = \min\biggl\{m: \forall v \in V, \norm{\rhobf_v - \pibf}_{TV} \leq 1/4 \biggr\}
\end{equation*}
where 
\begin{equation}
\label{eqn: uniform_random_walk}
\rhobf_{v}^{(m)} = \frac{1}{m}\sum_{k = 1}^{m+1} \qbf_{v}^{m}, ~~ \rhobf_v = \left( \rhobf_{v}^{(1)}, \rhobf_{v}^{(2)}, \rhobf_{v}^{(3)} \ldots \right), ~~ \rhobf = \left( \rhobf_v \right)_{v \in V}
\end{equation}
and $\norm{\pbf - \pibf}_{TV} = \sum_{v \in V}\abs{p_v - \pi_v}$ is the total variation norm between distributions $\pbf$ and $\pibf$. 
\begin{lemma}
	\label{lem: tv_mixing_to_pointwise_mixing1}
	For $\qbf$ as in \eqref{eqn: random_walk} and $\rhobf$ as in \eqref{eqn: uniform_random_walk},
	\begin{equation*}
	\tau_{\infty}(\qbf; G) \leq 2752 \tau_1(\rhobf; G) \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\}\right)
	\end{equation*}
\end{lemma}
\begin{proof}
	Masking dependence on the starting vertex $v$ for the moment, let
	\begin{equation*}
	\Delta_u^{(m)} = q_{vu}^{(m)} - \pi_u, ~~ \delta_u^{(m)} = \frac{\Delta_u^{(m)}}{\pi_u}
	\end{equation*}
	and $\Deltabf^{(m)} = (\Delta_u^{(m)})_{u \in V}$, $\deltabf^{(m)} = (\delta_u^{(m)})_{u \in V}$. For a vector $\Deltabf = (\Delta_u)_{u \in V}$, the $L^{p}(\pibf)$ norm is given by
	\begin{equation*}
	\norm{\Deltabf}_{L^p(\pibf)} = \left(\sum_{u \in V} \left(\Delta_u\right)^{p} \pi_u \right)^{1/p}
	\end{equation*}
	To go between the $L^{\infty}(\pibf)$ and $L^{1}(\pibf)$ norms, we have
	\begin{align*}
	\norm{\deltabf^{(2m)}}_{L^{\infty}(\pi)} & \overset{(i)}{\leq} \norm{\deltabf^{(m)}}^2_{L^{2}(\pi)} \\
	& = \norm{(\deltabf^{(m)})^2}_{L^{1}(\pi)} \\
	& \overset{(ii)}{\leq}  \norm{(\deltabf^{(m)})}_{L^{1}(\pi)} \norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)}
	\end{align*}
	where $(i)$ is a result of \cite{benjamini2003} and $(ii)$ follows from Holder's inequality. Now, we upper bound the second factor on the right hand side by observing
	\begin{align*}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} & \leq \max\left\{1, \max_{u \in V} \frac{q_{vu}^{(m)}}{\pi_u} \right\} \\
	& \overset{(iii)}{\leq} \max\left\{1, \frac{1}{s(G)}\right\}
	\end{align*}
	where $(iii)$ follows from Lemma \ref{lem: local_spread_G}.

	Now, we leverage the following well-known fact \cite{montenegro2002}: for any $\epsilon > 0$, if $m \geq \tau_1(\qbf_v^{(m)}; G) \cdot \log(1/\epsilon)$ then
	\begin{equation*}
	\norm{\qbf_v^{(m)} - \pibf}_{TV} \leq \epsilon.
	\end{equation*}
	But $\norm{\qbf_v^{(m)} - \pibf}_{TV}$ is exactly $\norm{(\deltabf^{(m)})}_{L^{1}(\pi)}$. Therefore, picking 
	\begin{equation*}
	m_0 = \tau_1(\qbf_v^{(m)}; G) \cdot \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\} \right)
	\end{equation*} implies $\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} \leq 1/4$ for all $m \geq 2 m_0$.  Then, 
	\begin{equation*}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} = \sup_{u}\left\{ \frac{\abs{q_{vu}^{(m)} - \pibf_u}}{\pibf_u} \right\}.
	\end{equation*}
	and since none of the above depended on a specific choice for $v$, the supremum can be taken over all starting vertices $v$ as well. Thus $\tau_{\infty}(\qbf^{(m)}; G) \leq 2m_0$.

	Finally, it is known \cite{montenegro2002} that
	\begin{equation*}
	\tau_{1}(\qbf^{(m)}; G) \leq 1376 \tau_{1}(\rhobf^{(m)}; G)
	\end{equation*}
	and so the desired result holds.	
\end{proof}

To prove Lemma \ref{lem: expansion_to_tv_distance} we first introduce a generalization of $\Phi(t;G) \cdot \Phi(t;G)$ known as a blocking conductance function.
\footnote{For more details, see \cite{montenegro2002}}

\begin{definition}[Blocking Conductance Function of \cite{montenegro2002}]
	\label{def: blocking_conductance_function}
	For $t_0 \geq \pi_{\min}$, a function $\phi(t; G): [t_0, 1/2] \to [0,1]$ is a \emph{blocking conductance function} if for all $S \subset V$ with $\pibf(S) = t \in [t_0, 1/2]$, either of the following hold: 
	\begin{enumerate}
		\item \emph{Exterior inequality.} For all $y \in \left[\frac{1}{2}t, t\right]: \phi_{int}(S) \geq \phi(\max\{t_0,y\})$
		\item \emph{Interior inequality.} For all $y  \in \left[t, \frac{3}{2}t\right]: \phi_{ext}(S) \geq \phi(\max\{y,1 - y\})$.
	\end{enumerate}
	where $\phi_{int}$ and $\phi_{ext}$ are defined respectively as
	\begin{align*}
	\phi_{int}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2} \\
	\phi_{ext}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S^c \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2}
	\end{align*}
\end{definition}

\begin{theorem}[Theorem 3.2 of \cite{montenegro2002}]
	\label{thm: montenegro}
	Consider $\phi(t; G): [t_0, 1/2] \to [0,1]$ a blocking conductance function. Then, letting
	\begin{equation*}
	h^m(t_0) = \sup_{S: \pibf(S) < t_0}  \bigl(\rhobf_v^{(m)}(S) - \pibf(S) \bigr)
	\end{equation*}
	the following statement holds: if $\phi$ is a blocking conductance function,
	\begin{align*}
	\norm{\rhobf_v^{(m)} - \pibf}_{TV} & \leq \max\left\{ \frac{1}{4}, h^1(t_0) +  \frac{70}{m}\left(\frac{1}{\phi(t_0; G)} + \int_{t = t_0}^{1/2} \frac{4}{t \phi(t; G)} dt\right) \right\} \\
	\end{align*}
\end{theorem}

Note that in \cite{montenegro2002} this theorem is stated with respect to $h^0$. However, in the subsequent proof it holds with respect to $h^m$, and it is observed that $h^m$ is decreasing in $m$. For our purposes it is more useful to state it with respect to $h^1$, as we have done.

\subsection{Conductance function and local spread: non-convex case.}
\label{sec: non_convex_conductance_function_and_local_spread}

Consider $\zbf \subset \Csig$ such that  $\mathcal{N}_{\zbf} = \set{B(z,r/3): z \in \zbf}$ is an internal covering of $\Csig$, meaning $\mathcal{N}_{\zbf} \supseteq \Csig$. Then, we write
\begin{align*}
\widetilde{B}_{\min} = \min_{z \in \zbf} \abs{B(z,r/3) \cap \wX}, ~~
\widetilde{D}_{\min} = \min_{\wx \in \wX} \abs{\widetilde{\cut}_{n,r}(x)} \\ \widetilde{B}_{\max} = \min_{z \in \zbf} \abs{B(z,r/3) \cap \wX}, ~~
\widetilde{D}_{\max} = \min_{\wx \in \wX} \abs{\widetilde{\cut}_{n,r}(x)}
\end{align*}
Both the conductance function and local spread will depend heavily on these quantities. Lemma \ref{lem: ball_bounds} collects the bounds we will need.
\begin{lemma}
	\label{lem: ball_bounds}
	Let $\Csig$ satisfy the conditions of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex}.	For sufficiently large $n$, and $r \leq \sigma/4d$, each of the following bounds hold with probability $1 - \delta$:
	\begin{align}
	\label{eqn: ball_bounds}
	\widetilde{B}_{\max} & \leq \left(1 + \sqrt{3^d\frac{3(\log \abs{\mathcal{N}_{\zbf}} + \log(1/\delta))}{n \nu_d r^d \Lambda_{\sigma}}}\right) n \nu_d \left(\frac{r}{3}\right)^d \Lambda_{\sigma} \nonumber \\
	\widetilde{B}_{\min} & \geq \left(1 - \sqrt{3^d\frac{2(\log \abs{\mathcal{N}_{\zbf}} + \log(1/\delta))}{n \nu_d r^d \lambda_{\sigma} \beta_d}}\right) n \nu_d \left(\frac{r}{3}\right)^d \lambda_{\sigma} \beta_d \nonumber \\
	\widetilde{D}_{\max} & \leq \left(1 + \sqrt{\frac{3(\log n + \log(1/\delta))}{n \nu_d r^d \Lambda_{\sigma}}}\right) n \nu_d r^d \Lambda_{\sigma} \nonumber \\
	\widetilde{D}_{\min} & \geq \left(1 - \sqrt{\frac{2(\log n + \log(1/\delta))}{n \nu_d r^d \lambda_{\sigma} \beta_d}}\right) n \nu_d \left(\frac{r}{3}\right)^d \lambda_{\sigma} \beta_d \nonumber \\
	\left(1 - \sqrt{\frac{2 \log(1/\delta)}{n \lambda_d \nu_d \sigma^d}}\right) n \lambda_{\sigma} \nu_d \sigma^d & \leq \widetilde{n} \leq \left(1 + \sqrt{\frac{3 \log(1/\delta)}{n \Lambda_{\sigma} \nu_d D^d}}\right) n \Lambda_d \nu_d D^d
	\end{align}
	where $\wn = \abs{\wX}$ and $\beta_d \geq 1/2^d$.

	In particular, fix $\epsilon > 0$. Then, for $n$ as specified in Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex}, the following event:
	\begin{align}
	\label{eqn: ball_bounds_epsilon}
	\widetilde{B}_{\max} \leq \left(1 + \epsilon \right) n \nu_d \left(\frac{r}{3}\right)^d \Lambda_{\sigma},& ~~ \widetilde{D}_{\max} \leq \left(1 + \epsilon \right) n \nu_d r^d \Lambda_{\sigma} \nonumber \\
	\widetilde{B}_{\min} \geq \left(1 - \epsilon \right) n \nu_d \left(\frac{r}{3}\right)^d \lambda_{\sigma} \beta_{d},& ~~ \widetilde{D}_{\min} \geq \left(1 - \epsilon \right) n \nu_d r^d \lambda_{\sigma} \beta_{d} \nonumber \\
	(1 - \epsilon) n \lambda_{\sigma} \nu_d \sigma^d & \leq \widetilde{n} \leq (1 + \epsilon) n \Lambda_{\sigma} \nu_d D^d 
	\end{align}
	occurs with probability at least $1 - \delta$.
\end{lemma}
\begin{proof}
	We observe that for any $s \leq \sigma/4d$ and any $x \in \Csig$,
	\begin{equation*}
	\nu(B(x,s) \cap \Csig) \geq \left(\frac{s}{2}\right)^d = s^d \beta_d.
	\end{equation*}
	(In fact, tighter bounds can be shown to hold, but we will not need them). Therefore, by \ref{asmp: bounded_density}:
	\begin{equation*}
	\lambda_{\sigma} \nu_d s^d \beta_d \leq \Pbb\left(B(z,s) \cap \Csig \right) \leq \Lambda_{\sigma} \nu_d s^d \\
	\end{equation*}
	In particular, this holds for $s = r$ and $s = r/3$, and for every $z \in \zbf$ as well as every $z \in \widetilde{\Xbf}$. Now, by \ref{asmp: cluster_diameter} and \ref{asmp: bounded_density} we also have
	\begin{equation*}
	\Lambda_{\sigma} \nu_d \sigma^d \leq \Pbb(\Csig) \leq \Lambda_{\sigma} \nu_d D^d
	\end{equation*}

	The proof of each statement in \eqref{eqn: ball_bounds} then follows from application of Lemma \ref{lem: multiplicative_Hoeffding}.

	To show \eqref{eqn: ball_bounds_epsilon}, we note that $\abs{\mathcal{N}_z}$ is less than the covering number of the $D$-ball in $d$ dimensions. Therefore
	\begin{equation*}
	\abs{\mathcal{N}_z} \leq \left(\frac{6D}{r} + 1\right)^d.
	\end{equation*}
	It is then immediately apparent that $n$ chosen as in Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex} yields \eqref{eqn: ball_bounds_epsilon}.
\end{proof}

Now, we consider the conductance function and local spread computed over $\widetilde{G}_{n,r}$, which we refer to by
\begin{equation*}
\widetilde{\Phi}_{n,r}(t) = \Phi(t; \widetilde{G}_{n,r}), ~~ \widetilde{s}_{n,r} = s(\widetilde{G}_{n,r}).
\end{equation*}
where the restriction in the minimization problem of \eqref{eqn: conductance_function} is with respect to $\widetilde{\pi}_{n,r}$ the stationary distribution over $\widetilde{G}_{n,r}$.

We will bound $\widetilde{\Phi}_{n,r}(1/2)$ and $\widetilde{s}_{n,r}$ under the event that \eqref{eqn: ball_bounds_epsilon} holds, noting that this occurs with probability at least $1 - \delta$.

\begin{lemma}
	\label{lem: local_spread_lower_bound}
	If the bounds given by \eqref{eqn: ball_bounds_epsilon} hold, then
	\begin{equation*}
	\widetilde{s}_{n,r} \geq \frac{9}{10} \frac{(1 + \epsilon)^2}{(1 - \epsilon)^2}\frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2} \frac{r^d}{D^d} \beta_d^2
	\end{equation*}
\end{lemma}

\begin{lemma}
	\label{lem: non_convex_conductance_lower_bound}
	If the bounds given by \eqref{eqn: ball_bounds_epsilon} hold, then:
	\begin{equation*}
	\widetilde{\Phi}_{n,r}(1/2) \geq \frac{\lambda_{\sigma}(1 - \epsilon)\beta_d}{4\Lambda_{\sigma}(1 + \epsilon)3^d} \left(1 + \frac{(1 - \epsilon)r^d\lambda_{\sigma}}{(1 + \epsilon)D^d\Lambda_{\sigma}}\right)
	\end{equation*}
\end{lemma}

	If, on the other hand , $\abs{\mathcal{N}_{\zbf}^+} = \mathcal{N}_{\zbf}$, then \eqref{eqn: normalized_cut_1} holds with respect to $S^c$. Then, because $\widetilde{\pi}_{n,r}(S) \leq 1/2$, 
	\begin{equation*}
	\frac{\widetilde{\cut}_{n,r}(S^c)}{\widetilde{\vol}_{n,r}(S^c)} \leq \frac{\widetilde{\cut}_{n,r}(S)}{\widetilde{\vol}_{n,r}(S)}
	\end{equation*}
	and so we get the exact statement of \eqref{eqn: normalized_cut_1}. Noting, as in the proof of Lemma \ref{lem: local_spread_lower_bound}, that  $\widetilde{\vol}_{n,r}(\widetilde{G}_{n,r}) \leq \wn \cdot \widetilde{D}_{\max}$, the relevant results of Lemma \ref{lem: ball_bounds} yield the desired inequality. 
\end{proof}

\subsection{Proof of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex}}
\label{sec: inverse_mixing_time_lower_bound_nonconvex}

Throughout this proof, we will condition on the events of Lemmas \ref{lem: local_spread_lower_bound} and \ref{lem: non_convex_conductance_lower_bound},
namely
\begin{align*}
\widetilde{s}_{n,r} & \geq \frac{9}{10} \frac{(1 + \epsilon)^2}{(1 - \epsilon)^2}\frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2} \frac{r^d}{D^d} \beta_d^2 \\
\widetilde{\Phi}_{n,r}(1/2) & \geq \frac{\lambda_{\sigma}(1 - \epsilon)\beta_d}{4\Lambda_{\sigma}(1 + \epsilon)3^d} \left(1 + \frac{(1 - \epsilon)r^d\lambda_{\sigma}}{(1 + \epsilon)D^d\Lambda_{\sigma}}\right)
\end{align*}
noting that for $n$ as chosen in Theorem \ref{thm: inverse_mixing_time_lower_bound}, this will occur with probability at least $1 - \delta$ (by Lemma \ref{lem: ball_bounds}).
